---
title: "The Great Doctors Visits Investigation"
author: "William Hopper"
date: "7/16/2020"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(moderndive)
```

## Matt's Data - Mercy Presbyterian
```{r}
mercy <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRzrU3hpegjpfOPjfxXpmlBVjPEQK5mb0Xz2Id2WtrAoUNh1ecIuGgYfWQic8M4-YIj87iKB8D_4hA2/pub?output=csv") %>%
  select(visits)%>%
  summarise(mean = mean(visits))
mercy
```

## Jon's Data - UNC Teaching Hospital
```{r}
UNC <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRd5i4J3Uns1vE_ehPyGj3RQI0bpxTVE5A4bzQ-EzrlOV0R9XGtIzPImMCbse3ykkMdiZ-sKCIoS7Mu/pub?output=csv") %>%
  select(visits)%>%
  summarise(mean = mean(visits))
UNC
```

## Christina's Data - Blossom Hill 
```{r}
blossom_hill <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQGXTBoreDm2-UUx81h56gbw8W5PU6-o8KfYyojPoG5LBEH-Bd7w39G_qRVNPaEOzPHHcnl7LRX7Ryj/pub?output=csv") %>%
  select(visits)%>%
  summarise(mean = mean(visits))
blossom_hill
```

Matt = 12.84715
Christina = 13.4
Jon = 13.80

```{r}
ncbirths_population <- read.csv("https://wjhopper.github.io/SDS-201/data/ncbirths_population.csv")
```

```{r}
birth_record_samples <- rep_sample_n(ncbirths_population,
                                     size = 5001,
                                     reps = 120
                                     )
avg_visits <- group_by(birth_record_samples, replicate) %>%
  summarise(mean_visits = mean(visits))
```

```{r}
ggplot(data=avg_visits, aes(x = mean_visits)) +
  geom_histogram() +
  scale_x_continuous(limits=c(12,15))
```


## Calculating the mean and variance of the sampling distribution

## Method 1: Simulation with 5001 records per sample
```{r}
summarise(avg_visits,
          mean(mean_visits),
          var(mean_visits)
          )
```

The mean of the sampling distribution was 13.1, and the variance of the sampling distribution was .0113.

## Method 2: Central Limit Theorem

```{r}
summarise(ncbirths_population,
          mean(visits),
          var(visits)
          )
```

The mean number of doctors visits in the population is 13.13, and the variance of the number of doctor's visits is 75.97

```{r}
mean_of_means = 13.13
variance_of_means = 75.97/5001
```

## Method 3: Central Limit Theorem ESTIMATION METHOD

Using the Mercy Hospital as data set to estimate the populatiom mean and variance.
```{r}
read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRzrU3hpegjpfOPjfxXpmlBVjPEQK5mb0Xz2Id2WtrAoUNh1ecIuGgYfWQic8M4-YIj87iKB8D_4hA2/pub?output=csv") %>%
  select(visits)%>%
  summarise(mean = mean(visits),
            variance = var(visits))
```

```{r}
estimated_mean_of_means <- 12.84
estimated_variance_of_means <- 27.91961/1001
```

