---
title: "Lab 1 Solutions: Intro to R and RStudio"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    toc_level: 2
    df_print: kable
    css: "../../assets/css/lab.css"
---

```{r include = F}
library(knitr) 
```

Please put all answers (both code chunks and text answers) below the "Solution" header for each exercise. 

<div id="license">
Remember to save your work as you go along! Click the save button in the upper left hand corner of the R Markdown window. 

It is a good idea to try kitting your document as you go along, to make sure your document is knitting, and that your html file includes Exercise headers, text, and code. Note that knitting automatically saves your Rmd file too!
</div>


## Exercise
First let's practice some basic formatting. Using <a href="https://rmarkdown.rstudio.com/authoring_basics.html" target="_blank">this formatting tips page</a>, figure out how to:
  
1. Italicize this sentence
2. Bold this sentence
3. Make the number two at the end of this sentence a superscript2

### Solution

1. *Italicize this sentence*
2. **Bold this sentence**
3. Make the number two at the end of this sentence a superscript^2

## Exercise
Answer the following with code in a code chunk (no text necessary). Remember that the code is just **instructions** for R. You need to run the code chunk to make R execute those instructions!

-   Create a variable called `y` with the value of 7 
-   Multiply `y` by `2.75`, and store the answer in a variable named `z`
    
### Solution
```{r exercise1}
y <- 7
z <- 2.5 * y
```


## Exercise
Run the following mathematical operation in a code chunk: `6 + 3`. Where does the answer appear? (please answer with **text**)

### Solution
```{r exercise2}
6 + 3
```

The answer 9 appears right below the code chunk in the R Markdown document if the "inline" output setting is on, or in the R console pane if the "console" output setting is on.

## Exercise
In a new code chunk, save the results of `6 + 3` as a variable called `a`. Does the answer appear? (please answer with **text**). Where does the object `a` show up? (please answer with **text**). Next type `a` into the code chunk and re-run the code chunk. What happens? (please answer with **text**)

### Solution
```{r exercise3}
a <- 6 + 3
a
```

The answer 9 does not appear below the code chunk. The object a shows up in the Environment pane, with a value of 9. When I added `a` into the code chunk and ran it, the answer 9 appeared


## Exercise
Run following command in a new code chunk. `a^2`. What does the `^` operator do? (please answer with **text**)

### Solution
```{r exercise4}
a^2
```

The `^` operator raises the variable to a power. In other words, it makes the 2 an exponent. In this case, we have squared the variable `a`, which has a value of 9, producing the value 81.

## Exercise
Type the following command into a new code chunk. `sum(a, 10, -1)`.`sum` is a function. Based on the output, what do you think the `sum` function does? (please answer with **text**)

### Solution
```{r exercise5}
sum(a, z, -1)
```

The sum function adds up all the values it is given, including adding variables and literal numeric values.

## Exercise
Click the little broom icon in the upper right hand corner of the **Environment** pane. Click yes on the window that opens. What happened? (please answer with **text**)

### Solution
The environment is now empty - the variables `a`, `y` and `z` have been deleted!

## Exercise
Go to the **Run** button at the top right of the R Markdown pane, and choose **Run All** (the last option). What happened? (please answer with **text**)

### Solution
All the code chunks in the document were executed again, and the variables `a`, `y` and `z` are present in the environment again.

## Exercise
Consider the following code chunk. What is the value of the variable `y` before running this code chunk. What is the value of the variable `y` after running this code chunk? Why? (please answer with **text**)

```{r eval = F}
y + 2
```

### Solution
The value of y is 7 befor and after running this code chunk. It is the same because we did not use the assignment operator `<-` to overwrite the original value of `y`.

## Exercise
Copy, paste, and run the following code to make a vector called `music`, that contains music genres. Recall a vector is a data object that has multiple elements of the same type. Here the data type is a **character**. Look in the environment pane. How does R tell us that this vector contains **characters**, not numbers? (please answer with **text**)

```{r}
music <- c("bluegrass", "funk", "folk")
```

### Solution
In in the Environment pane, the vector `music` is labeled as `chr`.


### Turning in your work
Once you are done, upload both this .Rmd file, and the HTML file you created by "knitting" this document to Moodle.

